---
title: "Final Project"
author: "Lucy Hu"
date: "April 26, 2020"
output: html_document
---

I started making some preliminary graphs for poverty and gender but I'm not sure if it would be better to display this as a grouped barplot or to display the overall rate as proportion of male vs female or some other display. 

Also started working through the poisson regression but am uncertain as to what values I should be inputting in the function. Based on the example from the notes, I found the  number of adolescents that were vaccinated (numMF) and the number of females that were vaccinated (numF) in the 13-17 age. However I can't seem to enter them in the function in a way that makes sense. 

I also added a new value that quantifies political leaning (called demAdv with a higher number indicating strong democratic leaning and negative numbers indicative republican leanings). However it seems that poisson regression cannot use negative numbers? Is there any meaningful way to incorporate this data? I'm very curious as the graph shows some interesting relationships. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(tidyr)
# any other libraries you want to load
```

### Question of interest

Human papillomavirus (HPV) is a group of related viruses that is often associated with being a common sexually transmitted infection. However, it has also been linked with various cancers such as cervical cancers, genital cancers, and oral/throat cancers. 

The HPV vaccine was first approved in 2006 and has been shown to be safe and effective at protecting against cancer-causing serotypes of the virus. It is also one of the most controversial vaccines and is only considered mandatory in 3 states and the capital of the US (District of Columbia, Hawaii, Rhode Island, and Virginia). In the District of Columbia, Hawaii, and Rhode Island, this policy applies to both boys and girls. However, in Virginia, the mandate only applies to girls. As such, vaccination rates across the US vary widely. 

The main question of interest in this analysis is: What factors are related to HPV vaccination rates?

### Data
The data for this analysis comes from the [CDC's 2018 Adolescent HPV Vaccination Coverage Report](https://www.cdc.gov/vaccines/imz-managers/coverage/teenvaxview/data-reports/hpv/reports/2018.html). This report contains coverage data from all 50 states as well as 3 US territories and the United States as a whole. It also contains data on race, poverty level, and number of vaccine doses. For the sake of this analysis, I'll be concentrating on the data for teens who are up to date with their HPV vaccination. 

I have also incorporated state party affiliation data for 2018, sourced from [Gallup polls](https://news.gallup.com/poll/247025/democratic-states-exceed-republican-states-four-2018.aspx).


<!-- Briefly describe the data you plan to use to answer the question of interest.  Provide links to any publically available datasets. You can directly link to websites where your data is by including links in the following way: Here in the link to the [Johns Hopkins University homepage](https://www.jhu.edu/). -->

```{r dataFull, include=FALSE}
## If you wanted to read in the full data set and name the columns in a useful way:
## for each rate, you want two columns: 2018 and SAMPLE SIZE
## you need to pull these columns, and then rename them in a useful way
## Right now, they have names like 2018, 2018_1, 2018_2...2018_35 and SAMPLE SIZE, SAMPLE SIZE_1, ... SAMPLE SIZE_35
HPV_data <- read_csv("HPV_CONDENSED.csv", skip=2)

HPV_rates <- HPV_data %>% select("Names", matches("2018"))
HPV_size <- HPV_data %>% select("Names", matches("SAMPLE SIZE"))


## from this, you can see that there are 12 different groupings of the variables Age, Poverty, Race/Ethnicity and Urbanicity: by Males and Females, Females alone, and Males alone
HPV_head1<- read_csv("HPV_CONDENSED.csv", n_max=1, col_names=FALSE)
# HPV_head1<-HPV_head1[1,][!is.na(HPV_head1[1,])]
HPV_head1 <- HPV_head1[,colSums(is.na(HPV_head1))<nrow(HPV_head1)]
## For example, you can get the first pieces using:
paste(rep(c("MF", "F", "M"), each=4), gsub(".*\\ ", "", HPV_head1), sep=".")

## from this, you can see that there are 36 total gender/variable groups
HPV_head2<- read_csv("HPV_CONDENSED.csv", skip=1, n_max=1, col_names=FALSE)
# HPV_head2<-HPV_head2[1,][!is.na(HPV_head2[1,])]
HPV_head2 <- HPV_head2[,colSums(is.na(HPV_head2))<nrow(HPV_head2)]

# remove spaces
gsub("\\ ", "", HPV_head2)

## create column names
## we want to appropriately replace column names to get, for example: MF.Age.13-17Years.Rate, MF.Poverty.AtOrAbove.Rate, MF.Poverty.Below....

colnames(HPV_rates)<- c("state", paste(rep(paste(rep(c("MF", "F", "M"), each=4), gsub(".*\\ ", "", HPV_head1), sep="."), times=rep(c(1,3,5,3), times=3)), gsub("\\ ", "", HPV_head2), "Rate", sep="."))

colnames(HPV_size)<- c("state", paste(rep(paste(rep(c("MF", "F", "M"), each=4), gsub(".*\\ ", "", HPV_head1), sep="."), times=rep(c(1,3,5,3), times=3)), gsub("\\ ", "", HPV_head2), "SampleSize", sep="."))

HPV_data <- HPV_rates %>% left_join(HPV_size)

```


```{r dataSimple, echo=FALSE}
## MAT: here is another version of the code to just pull the 13-17 M/F results. You will want the sample size for modeling.


HPV_data = HPV_data[-c(52,53,54,55),]
# HPV_data<- HPV_data %>% select(state=Names, MF.Age.13to17Years.Rate = `2018`, MF.Age.13to17Years.SampleSize=`SAMPLE SIZE`)
# HPV_data <- HPV_data %>% mutate(state=toupper(state))


politics_data <- read_csv("leanings detailed.csv", col_names=FALSE)
politics_data <- politics_data %>% select(state=X1, leaning=X3, demAdv=X2) 

HPV_data <- HPV_data %>% left_join(politics_data, by="state")

```

### Variables of interest

* **Outcome variable: State HPV vaccine coverage**
* **Primary predictor variable: poverty, gender**
* **Possible confounders: political leaning**
* **Potential effect modifiers:**

### Directed acyclic graph (DAG)


### Analysis plan
I will be using poisson regression to analyze this dataset because it is count data. 

<!-- What type of basic analysis could you use to address your question of interest?  Linear regression because your outcome is continuous?  Logistic regression because your outcome is binary?  How will you control for confounding in your analysis?  This section may change and be updated later as you continue with your analysis. -->

### Preliminary graphs 
To get an overview of what the current status of vaccination is in the United States and some territories, this is a plot of the percentage of teens (ages 13-17) who are up to date with their HPV vaccines. 
```{r, echo=FALSE}
MFagePlot <- ggplot(HPV_data, aes(x=state, y=`MF.Age.13-17Years.Rate`)) +
  geom_bar(stat="identity",  fill="steelblue") +
  theme_minimal() + 
  geom_text(aes(label=`MF.Age.13-17Years.Rate`), vjust=-0.3, size=2) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") 

MFagePlot
```

I've also presented this information sorted by rate:
```{r, echo=FALSE}
HPV_data$stateSort <- factor(HPV_data$state, levels = HPV_data$state[order(HPV_data$`MF.Age.13-17Years.Rate`)])

# palette
pal <- c(
  "Strong Democratic" = "#3449eb",
  "Strong Republican" = "#eb3434",
  "Competitive" = "#c5c6c9",
  "Lean Democratic" = "#7abafa",
  "Lean Republican" = "#fa7a83"
)

MFsortedPlot <- ggplot(HPV_data, aes(x=stateSort, y=`MF.Age.13-17Years.Rate`)) +
  geom_bar(stat="identity",  fill="steelblue") +
  theme_minimal() + 
  geom_text(aes(label=`MF.Age.13-17Years.Rate`), vjust=-0.3, size=2) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") + 
  scale_fill_manual(
    values = pal,
    limits = names(pal)
  )

MFsortedPoliticsPlot <- ggplot(HPV_data, aes(x=stateSort, y=`MF.Age.13-17Years.Rate`, fill=leaning)) +
  geom_bar(stat="identity") +
  theme_minimal() + 
  geom_text(aes(label=`MF.Age.13-17Years.Rate`), vjust=-0.3, size=2.5) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") + 
  scale_fill_manual(
    values = pal,
    limits = names(pal)
  )

MFsortedPlot
MFsortedPoliticsPlot
```

I also looked at the vaccination rates for males vs females for each state:

```{r, echo=FALSE}
# male rates()
male <- HPV_data %>%
  transmute(state, 
            rate = `M.Age.13-17Years.Rate`, 
            gender = "Male")
# female rates
female <- HPV_data %>%
  transmute(state, 
            rate = `F.Age.13-17Years.Rate`, 
            gender = "Female") 
gender_rates <- rbind(male, female)

# HPV_data %>%
#   count(`F.Age.13-17Years.Rate`,`M.Age.13-17Years.Rate`) %>%
#   group_by() %>%
#   mutate(prop = n/sum(n), N=sum(n)) %>%
#   ungroup()

genderPlot <- ggplot(gender_rates) + 
  geom_bar(aes(x=state, y=rate, fill=gender), stat="identity", position="dodge") +
  theme_minimal() + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State and Gender", x = "State", y = "Rate") +
  scale_y_continuous("Rate", expand = c(0, 0)) +
  scale_x_discrete("State") 
  # scale_fill_manual(values = c("#468189", "#9DBEBB"))

genderPlot
```



Next, I plotted rates for those above/below the poverty line, removing states with missing data:
```{r}
# above poverty rates()
notPoor <- HPV_data %>%
  transmute(state, 
            rate = `MF.Poverty.LivingAtorAbovePoverty.Rate`, 
            status = "Not poor")
# below poverty rates
poor <- HPV_data %>%
  transmute(state, 
            rate = `MF.Poverty.LivingBelowPoverty.Rate`, 
            status = "Poor") 
poverty_rates <- rbind(notPoor, poor)


# select only states with complete data
poverty_rates <- drop_na(poverty_rates)

# HPV_data %>%
#   count(`F.Age.13-17Years.Rate`,`M.Age.13-17Years.Rate`) %>%
#   group_by() %>%
#   mutate(prop = n/sum(n), N=sum(n)) %>%
#   ungroup()

povertyPlot <- ggplot(poverty_rates) + 
  geom_bar(aes(x=state, y=rate, fill=status), stat="identity", position="dodge") +
  theme_minimal() + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State and Poverty Status", x = "State", y = "Rate") +
  scale_y_continuous("Rate", expand = c(0, 0)) +
  scale_x_discrete("State") 
  # scale_fill_manual(values = c("#468189", "#9DBEBB"))

povertyPlot
```

Plotting overall rates by gender
```{r}
### Code for plotting overall rate for males/females
overall_gender_data <- HPV_data %>%
  mutate(numMF = `MF.Age.13-17Years.Rate`*`MF.Age.13-17Years.SampleSize`,
         numF = `F.Age.13-17Years.Rate`*`F.Age.13-17Years.SampleSize`,
         numM = `M.Age.13-17Years.Rate`*`M.Age.13-17Years.SampleSize`) %>%
  summarize(rate_overall = sum(numMF)/sum(`MF.Age.13-17Years.SampleSize`),
            rate_F = sum(numF)/sum(`F.Age.13-17Years.SampleSize`),
            rate_M = sum(numM)/sum(`M.Age.13-17Years.SampleSize`))
overall_gender_data

overall_genderPlot <- ggplot(overall_gender_data) + 
  geom_bar(aes(x = "Overall", y = rate_overall), stat="identity", position="dodge") +
  geom_bar(aes(x = "Female", y = rate_F), stat="identity", position="dodge", fill="red") + 
  geom_bar(aes(x = "Male", y = rate_M), stat="identity", position="dodge", fill="blue") +
  theme_minimal() + 
  labs(title = "HPV Vaccination Rates Overall and by Gender", x = "", y = "Rate") +
  scale_y_continuous("Rate", expand = c(0, 0))
  # scale_fill_manual(values = c("#468189", "#9DBEBB"))

overall_genderPlot
```

### Analysis
```{r}
HPV_data <- HPV_data %>% mutate(numMF = round(`MF.Age.13-17Years.Rate`*`MF.Age.13-17Years.SampleSize`),
                                numF = round(`F.Age.13-17Years.Rate`*`F.Age.13-17Years.SampleSize`),
                                numM = round(`M.Age.13-17Years.Rate`*`M.Age.13-17Years.SampleSize`),
                                ssMF = `MF.Age.13-17Years.SampleSize`,
                                ssF = `F.Age.13-17Years.SampleSize`,
                                ssM = `M.Age.13-17Years.SampleSize`)

# Shows the effect of political leaning on the overall rates
# The reference group is "Competitive"
# So we see that overall states that lean or are strong Dem have about a 20% increase in the vaccine rate compared to competitive states
# States that lean or are strong Republican have about a 10% decrease in the vaccine rate compared to competitive states
model_1 = glm(numMF ~ leaning, data=HPV_data, offset=log(ssMF), family=poisson(link="log"))
summary(model_1)
exp(model_1$coefficients)
exp(confint.default(model_1))


# Shows the effect of political leaning on the rates for females
# The reference group is "Competitive"
# For female vaccination rates, it looks like the effects are a little less strong
model_2 = glm(numF ~ leaning, data=HPV_data, offset=log(ssF), family=poisson(link="log"))
summary(model_2)
exp(model_2$coefficients)
exp(confint.default(model_2))


# Shows the effect of political leaning on the rates for males
# The reference group is "Competitive"
# For males vaccination rates, it looks like the effects are a little stronger!
model_3 = glm(numM ~ leaning, data=HPV_data, offset=log(ssM), family=poisson(link="log"))
summary(model_3)
exp(model_3$coefficients)
exp(confint.default(model_3))


```




```{r submission_instructions, include=FALSE, eval=FALSE}
REMINDER:  When you are ready to submit your assignment, do ALL of the following:

* First, knit your .Rmd file to see if it will compile.  If it doesn't knit, you can still follow the steps below to submit your assignment, but please try to resolve the knitting issues before you submit.  You can reach out to us at phbiostats@jhu.edu for help!
* Next you need to **commit** the changes you've made to the document.  Click the colorful Git button at the top of the RStudio window and select "Commit" from the menu.
* In the window that opens, **stage** your changes by clicking the check boxes next to the Rmd file.
* In the "Commit message" box, type a short description of what changes you've made, something like: `Completed assignment`
* Click the "Commit" button on the bottom right.
* You'll see a window that will tell you about the changes that you've made.  Click "Close" to close the window.  You've successfully committed! You can close this commit window now.
* After committing, you must **push** your changes to the repository on Github.  Do this by clicking the colorful Git button again and select "Push Branch".  
* Again, you'll see a window open that tells you your changes have been pushed!
* If you want, you can look at your repository on [Github.com](https://github.com/) and should be able to see your changes there!  
* You've successfully submitted your assignment :)
```
