---
title: "Final Project"
author: "Lucy Hu"
date: "April 12, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
# any other libraries you want to load
```

### Question of interest

Human papillomavirus (HPV) is a group of related viruses that is often associated with being a common sexually transmitted infection. However, it has also been linked with various cancers such as cervical cancers, genital cancers, and oral/throat cancers. 

The HPV vaccine was first approved in 2006 and has been shown to be safe and effective at protecting against cancer-causing serotypes of the virus. It is also one of the most controversial vaccines and is only considered mandatory in 4 regions of the US (Washington DC, Hawaii, Rhode Island, and Virginia). As such, vaccination rates across the US vary widely. 

The main question of interest in this analysis is: What factors are related to HPV vaccination rates?

### Data
The data for this analysis comes from the [CDC's 2018 Adolescent HPV Vaccination Coverage Report](https://www.cdc.gov/vaccines/imz-managers/coverage/teenvaxview/data-reports/hpv/reports/2018.html). This report contains coverage data from all 50 states as well as 3 US territories and the United States as a whole. It also contains data on race, poverty level, and number of vaccine doses. For the sake of this analysis, I'll be concentrating on the data for teens who are up to date for the HPV vaccine. 


<!-- Briefly describe the data you plan to use to answer the question of interest.  Provide links to any publically available datasets. You can directly link to websites where your data is by including links in the following way: Here in the link to the [Johns Hopkins University homepage](https://www.jhu.edu/). -->

```{r data, include=FALSE}
# eventually you will include code to load your data here
HPV_data <- read_csv("HPV_CONDENSED.csv")
politics_data <- read_csv("leanings 2.csv")

df <- data.frame(HPV_data)
p_df <- data.frame(politics_data)
p2_df <- p_df[ , "leanings", drop = FALSE]



# <- dataframe[c(row numbers), c(column numbers)], numbers are inclusive
named_df <- 
  df %>% 
    select(states = X1,
           rates = HPV.Vaccination.Up.To.Date..Males.and.Females...Age
           )
overview_df <- named_df[c(3:54),c(1:2)]

overview_df$leanings=p2_df

order <- order(overview_df$rates)
sorted <- overview_df[order, ]


sorted <- sorted %>%
  mutate (leanings = factor(leanings,
                            level = c("blue","red","gray"),
                            labels = c("blue","red","gray"))
          )

```

### Variables of interest

* **Outcome variable: State HPV vaccine coverage**
* **Primary predictor variable: poverty**
* **Possible confounders: political leaning, gender**
* **Potential effect modifiers:**

### Directed acyclic graph (DAG)


### Analysis plan
Most likely will use linear regression as this outcome is not binary. I will probably need to find additional data and information on population characteristics in order to control for  confounding. 

<!-- What type of basic analysis could you use to address your question of interest?  Linear regression because your outcome is continuous?  Logistic regression because your outcome is binary?  How will you control for confounding in your analysis?  This section may change and be updated later as you continue with your analysis. -->

### Preliminary graphs 
To get an overview of what the current status of vaccination is in the United States and some territories, this is a plot of the percentage of teens (ages 13-17) who are up to date with their HPV vaccines. 
```{r, echo=FALSE}
ggplot(overview_df, aes(x=states, y=rates)) +
  geom_bar(stat="identity",  fill="steelblue") +
  theme_minimal() + 
  geom_text(aes(label=rates), vjust=-0.3, size=2) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") 
```

I've also presented this information sorted by rate:
```{r, echo=FALSE}
sorted$states <- factor(sorted$states, levels = sorted$states[order(sorted$rates)])

# palette
pal <- c(
  "blue" = "#3449eb",
  "red" = "#eb3434",
  "gray" = "#c5c6c9"
)

p0 <- ggplot(sorted, aes(x=states, y=rates)) +
  geom_bar(stat="identity",  fill="steelblue") +
  theme_minimal() + 
  geom_text(aes(label=rates), vjust=-0.3, size=2) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") + 
  scale_fill_manual(
    values = pal,
    limits = names(pal)
  )

p <- ggplot(sorted, aes(x=states, y=rates, fill="leanings")) +
  geom_bar(stat="identity") +
  theme_minimal() + 
  geom_text(aes(label=rates), vjust=-0.3, size=3) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") + 
  scale_fill_manual(
    values = pal,
    limits = names(pal)
  )

p0
p
```


```{r submission_instructions, include=FALSE, eval=FALSE}
REMINDER:  When you are ready to submit your assignment, do ALL of the following:

* First, knit your .Rmd file to see if it will compile.  If it doesn't knit, you can still follow the steps below to submit your assignment, but please try to resolve the knitting issues before you submit.  You can reach out to us at phbiostats@jhu.edu for help!
* Next you need to **commit** the changes you've made to the document.  Click the colorful Git button at the top of the RStudio window and select "Commit" from the menu.
* In the window that opens, **stage** your changes by clicking the check boxes next to the Rmd file.
* In the "Commit message" box, type a short description of what changes you've made, something like: `Completed assignment`
* Click the "Commit" button on the bottom right.
* You'll see a window that will tell you about the changes that you've made.  Click "Close" to close the window.  You've successfully committed! You can close this commit window now.
* After committing, you must **push** your changes to the repository on Github.  Do this by clicking the colorful Git button again and select "Push Branch".  
* Again, you'll see a window open that tells you your changes have been pushed!
* If you want, you can look at your repository on [Github.com](https://github.com/) and should be able to see your changes there!  
* You've successfully submitted your assignment :)
```
