---
title: "Final Project"
author: "Lucy Hu"
date: "April 12, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
# any other libraries you want to load
```

### Question of interest

Human papillomavirus (HPV) is a group of related viruses that is often associated with being a common sexually transmitted infection. However, it has also been linked with various cancers such as cervical cancers, genital cancers, and oral/throat cancers. 

The HPV vaccine was first approved in 2006 and has been shown to be safe and effective at protecting against cancer-causing serotypes of the virus. It is also one of the most controversial vaccines and is only considered mandatory in 4 regions of the US (Washington DC, Hawaii, Rhode Island, and Virginia). As such, vaccination rates across the US vary widely. 

The main question of interest in this analysis is: What factors are related to HPV vaccination rates?

### Data
The data for this analysis comes from the [CDC's 2018 Adolescent HPV Vaccination Coverage Report](https://www.cdc.gov/vaccines/imz-managers/coverage/teenvaxview/data-reports/hpv/reports/2018.html). This report contains coverage data from all 50 states as well as 3 US territories and the United States as a whole. It also contains data on race, poverty level, and number of vaccine doses. For the sake of this analysis, I'll be concentrating on the data for teens who are up to date for the HPV vaccine. 


<!-- Briefly describe the data you plan to use to answer the question of interest.  Provide links to any publically available datasets. You can directly link to websites where your data is by including links in the following way: Here in the link to the [Johns Hopkins University homepage](https://www.jhu.edu/). -->

```{r dataFull, include=FALSE, eval=FALSE}
## MAT: I didn't read carefully that you just wanted to focus on the 13-17 year old group, so I wrote a bunch of code that works to create a somewhat nicely labeled version of the full set of data. Since it sounds like you don't need it, I'll just leave it here, but add code down below that does what you want it to. I've added an eval=FALSE tag above to help.

## If you wanted to read in the full data set and name the columns in a useful way:
## for each rate, you want two columns: 2018 and SAMPLE SIZE
## you need to pull these columns, and then rename them in a useful way
## Right now, they have names like 2018, 2018_1, 2018_2...2018_35 and SAMPLE SIZE, SAMPLE SIZE_1, ... SAMPLE SIZE_35
HPV_data <- read_csv("HPV_CONDENSED.csv", skip=2)
HPV_data<- HPV_data %>% select("Names", contains(c("2018", "SAMPLE SIZE")))

## from this, you can see that there are 12 different groupings of the variables Age, Poverty, Race/Ethnicity and Urbanicity: by Males and Females, Females alone, and Males alone
HPV_head1<- read_csv("HPV_CONDENSED.csv", n_max=1, col_names=FALSE)
HPV_head1<-HPV_head1[1,][!is.na(HPV_head1[1,])]
## For example, you can get the first pieces using:
paste(rep(c("MF", "F", "M"), each=4), gsub(".*\\ ", "", HPV_head1), sep=".")

## from this, you can see that there are 36 total gender/variable groups
HPV_head2<- read_csv("HPV_CONDENSED.csv", skip=1, n_max=1, col_names=FALSE)
HPV_head2<-HPV_head2[1,][!is.na(HPV_head2[1,])]
# remove spaces
gsub("\\ ", "", HPV_head2)

## create column names
## we want to appropriately replace column names to get, for example: MF.Age.13-17Years.Rate, MF.Poverty.AtOrAbove.Rate, MF.Poverty.Below....
## here is what they will look like
c("state", paste(rep(paste(rep(c("MF", "F", "M"), each=4), gsub(".*\\ ", "", HPV_head1), sep="."), times=rep(c(1,3,5,3), times=3)), gsub("\\ ", "", HPV_head2), "Rate", sep="."), paste(rep(paste(rep(c("MF", "F", "M"), each=4), gsub(".*\\ ", "", HPV_head1), sep="."), times=rep(c(1,3,5,3), times=3)), gsub("\\ ", "", HPV_head2), "SampleSize", sep="."))

colnames(HPV_data)<- c("state", paste(rep(paste(rep(c("MF", "F", "M"), each=4), gsub(".*\\ ", "", HPV_head1), sep="."), times=rep(c(1,3,5,3), times=3)), gsub("\\ ", "", HPV_head2), "Rate", sep="."), paste(rep(paste(rep(c("MF", "F", "M"), each=4), gsub(".*\\ ", "", HPV_head1), sep="."), times=rep(c(1,3,5,3), times=3)), gsub("\\ ", "", HPV_head2), "SampleSize", sep="."))

```


```{r dataSimple, include=FALSE}
## MAT: here is another version of the code to just pull the 13-17 M/F results. You will want the sample size for modeling.

HPV_data <- read_csv("HPV_CONDENSED.csv", skip=2)
HPV_data<- HPV_data %>% select(state=Names, MF.Age.13to17Years.Rate = `2018`, MF.Age.13to17Years.SampleSize=`SAMPLE SIZE`)
HPV_data <- HPV_data %>% mutate(state=toupper(state))


## MAT: I would suggest that you want to keep the state names when you read in the leanings
## MAT: you want to read in the version with the state names, to be sure that they match when you merge the two data sets
politics_data <- read_csv("leanings.csv", col_names=FALSE)
politics_data <- politics_data %>% select(state=X1, leaning=X2) %>% mutate(state=toupper(state))

## note that some of the state names do not match between the two data sets, e.g., the leanings one has VA instead of VIRGINIA in the HPV data set; this will need to be cleaned up
## you will see that there are bars with missing data in the plot below

## MAT: using the left_join function here will match on the values in the state variable, to make sure
## you are merging things correctly
HPV_data <- HPV_data %>% left_join(politics_data)

```

### Variables of interest

* **Outcome variable: State HPV vaccine coverage**
* **Primary predictor variable: poverty**
* **Possible confounders: political leaning, gender**
* **Potential effect modifiers:**

### Directed acyclic graph (DAG)


### Analysis plan
Most likely will use linear regression as this outcome is not binary. I will probably need to find additional data and information on population characteristics in order to control for  confounding. 

<!-- What type of basic analysis could you use to address your question of interest?  Linear regression because your outcome is continuous?  Logistic regression because your outcome is binary?  How will you control for confounding in your analysis?  This section may change and be updated later as you continue with your analysis. -->

### Preliminary graphs 
To get an overview of what the current status of vaccination is in the United States and some territories, this is a plot of the percentage of teens (ages 13-17) who are up to date with their HPV vaccines. 
```{r, echo=FALSE}
ggplot(HPV_data, aes(x=state, y=MF.Age.13to17Years.Rate)) +
  geom_bar(stat="identity",  fill="steelblue") +
  theme_minimal() + 
  geom_text(aes(label=MF.Age.13to17Years.Rate), vjust=-0.3, size=2) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") 
```

I've also presented this information sorted by rate:
```{r, echo=FALSE}
HPV_data$stateSort <- factor(HPV_data$state, levels = HPV_data$state[order(HPV_data$MF.Age.13to17Years.Rate)])

# palette
pal <- c(
  "blue" = "#3449eb",
  "red" = "#eb3434",
  "gray" = "#c5c6c9"
)

p0 <- ggplot(HPV_data, aes(x=stateSort, y=MF.Age.13to17Years.Rate)) +
  geom_bar(stat="identity",  fill="steelblue") +
  theme_minimal() + 
  geom_text(aes(label=MF.Age.13to17Years.Rate), vjust=-0.3, size=2) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") + 
  scale_fill_manual(
    values = pal,
    limits = names(pal)
  )

p <- ggplot(HPV_data, aes(x=stateSort, y=MF.Age.13to17Years.Rate, fill=leaning)) +
  geom_bar(stat="identity") +
  theme_minimal() + 
  geom_text(aes(label=MF.Age.13to17Years.Rate), vjust=-0.3, size=3) + 
  theme(axis.text.x = element_text(size=9, angle=90),
        axis.text.y = element_text(size=9)) +
  labs(title = "HPV Vaccination Rates by State", x = "State", y = "Rate") + 
  scale_fill_manual(
    values = pal,
    limits = names(pal)
  )

p0
p
```


```{r submission_instructions, include=FALSE, eval=FALSE}
REMINDER:  When you are ready to submit your assignment, do ALL of the following:

* First, knit your .Rmd file to see if it will compile.  If it doesn't knit, you can still follow the steps below to submit your assignment, but please try to resolve the knitting issues before you submit.  You can reach out to us at phbiostats@jhu.edu for help!
* Next you need to **commit** the changes you've made to the document.  Click the colorful Git button at the top of the RStudio window and select "Commit" from the menu.
* In the window that opens, **stage** your changes by clicking the check boxes next to the Rmd file.
* In the "Commit message" box, type a short description of what changes you've made, something like: `Completed assignment`
* Click the "Commit" button on the bottom right.
* You'll see a window that will tell you about the changes that you've made.  Click "Close" to close the window.  You've successfully committed! You can close this commit window now.
* After committing, you must **push** your changes to the repository on Github.  Do this by clicking the colorful Git button again and select "Push Branch".  
* Again, you'll see a window open that tells you your changes have been pushed!
* If you want, you can look at your repository on [Github.com](https://github.com/) and should be able to see your changes there!  
* You've successfully submitted your assignment :)
```
